---
title: "TutorialImputingMissingData"
output: pdf_document
---

```{r, global_options, echo=FALSE}
# Clean workspace
rm(list=ls())

suppressWarnings(library(knitr)) # quiet starting knitr library

## sets the root directory for code chunks to better organize the files in Rstudio
# root.dir: (NULL) the root directory when evaluating code chunks; if NULL, the 
# directory of the input document will be used:
opts_knit$set(root.dir=normalizePath('../'))

# set options for figures
opts_chunk$set(fig.width=6, fig.height=4, fig.path='../Figures/',
            echo=TRUE, warning=FALSE, message=FALSE)

# set global controll flag to selectively run chunks 
eval_everything=FALSE
```

# Tutorial on Missing Data Imputation

We follow the tutorial on R packages for missing data imputation by  MANISH SARASWAT which can be found [here](https://www.analyticsvidhya.com/blog/2016/03/tutorial-powerful-packages-imputing-missing-values/), but discuss the mi package first:

## mi

[mi](http://www.stat.columbia.edu/~gelman/research/published/mipaper.pdf) is the package by Dr. Gelman, maintained bt Dr. Goodrich and uses predictive mean matching (pmm). predictive mean matching is explained well [here](https://www.google.com "Google's Homepage"), by  Paul Allsion on Statistical Horizon with statistical background and pitfalls, referencing also [(Morris 2014)](http://bmcmedresmethodol.biomedcentral.com/articles/10.1186/1471-2288-14-75).




```{r libraries_mi}
library(mi)
library(missForest)
```

```{r seed_missing_data_mi}
data("iris")

# seed missing values ( 10% )
iris.mis <- prodNA(iris, noNA = 0.1)
summary(iris.mis)

# imputing missing value with mi
mi_data <- mi(iris.mis, seed = 335)
summary(mi_data)
``` 

## MICE Package

```{r libraries_MICE}
library(missForest)
library(mice)
library(VIM)
```

```{r data, echo=TRUE}
data <- iris
```

### Generate Missing Data with **missForest**

Generate 10% missing values at Random using the missForest package

```{r}
iris.mis <- prodNA(iris, noNA = 0.1)
summary(iris.mis)
```

### Remove categorical variables and focus on continuous variables

```{r}
iris.mis <- subset(iris.mis, select = -c(Species))
summary(iris.mis)
```

### Inspect Missing Pattern with MICE

```{r}
md.pattern(iris.mis)
md.pattern(iris.mis)
```

### Visual Inspection of Missing Patern with VIM

```{r}
mice_plot <- aggr(iris.mis, col=c('navyblue','yellow'),
                    numbers=TRUE, sortVars=TRUE,
                    labels=names(iris.mis), cex.axis=.7,
                    gap=3, ylab=c("Missing data","Pattern"))
```

### Imputing the missing data with MICE

```{r impute_MICE, results='hide'}
imputed_Data <- mice(iris.mis, m=5, maxit = 50, method = 'pmm', seed = 500)
```

```{r}
summary(imputed_Data)
#check imputed values
imputed_Data$imp$Sepal.Width
#get complete data ( 2nd out of 5)
completeData <- complete(imputed_Data,2)
```

### Build a model using the imputed data

```{r}
#build predictive model
#Caveat I deviate from the Tutorial by using imputed_Data instead of iris.mis, because it otherwise throughs an error
fit <- with(data = imputed_Data, exp = lm(Sepal.Width ~ Sepal.Length + Petal.Width)) 

#combine results of all 5 models
combine <- pool(fit)
summary(combine)
```

### Build a model without imputation to compare

```{r}
raw.data <- iris
poor_fit <- fit <- with(data = raw.data, exp = lm(Sepal.Width ~ Sepal.Length + Petal.Width)) 
summary(poor_fit)
```

The point estimates of the poor_fit regression summary (without imputation) differ from the regression coefficients based on the imputed data; the latter also have wider confidence bands expressing the increased uncertainty due to imputation.

## AMELIA package


```{r libraries_Amelia}
rm(list = setdiff(ls(), lsf.str())) # uses setdiff to identify all  objects that are NOT functions.
# and remoes them
# lsf.str() finds all functions 

library(Amelia)
library(missForest)
```


## Seed 10% missing values
```{r generate_missing_Amelia}
data("iris")
iris.mis <- prodNA(iris, noNA = 0.1)
summary(iris.mis)
```

## Specify columns and run amelia
```{r run_Amelia}
amelia_fit <- amelia(iris.mis, m=5, parallel = "multicore", noms = "Species")
# access imputed outputs
amelia_fit$imputations[[1]]
# ...
```

## missForest package

```{r missingForest}
#missForest
library(missForest)

#load data
data("iris")

#seed 10% missing values
iris.mis <- prodNA(iris, noNA = 0.1)
summary(iris.mis)
```

## Hmisc
another package discussed but which I did not explore so far.
