---
title: "TutorialImputingMissingData"
output: pdf_document
---

```{r, global_options, echo=FALSE}
# Clean workspace
rm(list=ls())

suppressWarnings(library(knitr)) # quiet starting knitr library

## sets the root directory for code chunks to better organize the files in Rstudio
# root.dir: (NULL) the root directory when evaluating code chunks; if NULL, the 
# directory of the input document will be used:
opts_knit$set(root.dir=normalizePath('../'))

# set options for figures
opts_chunk$set(fig.width=6, fig.height=4, fig.path='../Figures/',
            echo=TRUE, warning=FALSE, message=FALSE)

# set global controll flag to selectively run chunks 
eval_everything=FALSE
```

# Tutorial on Missing Data Imputation

We follow the tutorial on R packages for missing data imputation by  MANISH SARASWAT which can be found here:
<https://www.analyticsvidhya.com/blog/2016/03/tutorial-powerful-packages-imputing-missing-values/>

## MICE Package

```{r libraries_MICE}
library(missForest)
library(mice)
library(VIM)
```

```{r data, echo=TRUE}
data <- iris
```

### Generate Missing Data with **missForest**

Generate 10% missing values at Random using the missForest package

```{r}
iris.mis <- prodNA(iris, noNA = 0.1)
summary(iris.mis)
```

### Remove categorical variables and focus on continuous variables

```{r}
iris.mis <- subset(iris.mis, select = -c(Species))
summary(iris.mis)
```

### Inspect Missing Pattern with MICE

```{r}
md.pattern(iris.mis)
md.pattern(iris.mis)
```

### Visual Inspection of Missing Patern with VIM

```{r}
mice_plot <- aggr(iris.mis, col=c('navyblue','yellow'),
                    numbers=TRUE, sortVars=TRUE,
                    labels=names(iris.mis), cex.axis=.7,
                    gap=3, ylab=c("Missing data","Pattern"))
```

### Imputing the missing data with MICE

```{r impute_MICE, results='hide'}
imputed_Data <- mice(iris.mis, m=5, maxit = 50, method = 'pmm', seed = 500)
```

```{r}
summary(imputed_Data)
#check imputed values
imputed_Data$imp$Sepal.Width
#get complete data ( 2nd out of 5)
completeData <- complete(imputed_Data,2)
```

### Build a model using the imputed data

```{r}
#build predictive model
#Caveat I deviate from the Tutorial by using imputed_Data instead of iris.mis, because it otherwise throughs an error
fit <- with(data = imputed_Data, exp = lm(Sepal.Width ~ Sepal.Length + Petal.Width)) 

#combine results of all 5 models
combine <- pool(fit)
summary(combine)
```

### Build a model without imputation to compare

```{r}
raw.data <- iris
poor_fit <- fit <- with(data = raw.data, exp = lm(Sepal.Width ~ Sepal.Length + Petal.Width)) 
summary(poor_fit)
```

The point estimates of the poor_fit regression summary (without imputation) differ from the regression coefficients based on the imputed data; the latter also have wider confidence bands expressing the increased uncertainty due to imputation.

## AMELIA package